%\documentclass[12pt,draftcls]{ucdavisthesis}
\documentclass[12pt, nofigureslist, notableslist,twoside]{ucdavisthesis}

% PLEASE READ THE MANUAL - ucdavisthesis.pdf (in the package installation directory)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                      %
%               LATEX COMMANDS FOR DOCUMENT SETUP                      %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\usepackage{bookmark}
\usepackage{lmodern}
\usepackage[us,nodayofweek,12hr]{datetime}
\usepackage{graphicx}
\usepackage{rotating}
\usepackage{subfigure}
\usepackage[none]{hyphenat}
\emergencystretch 3em
\usepackage[hyphens]{url}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{enumitem}
\usepackage[capitalize]{cleveref}
\usepackage{multirow}
\usepackage{comment}
\usepackage{amsfonts}
\usepackage{courier}
\usepackage{amssymb}
\usepackage{rotating}
\usepackage{afterpage}
\usepackage{mathtools}
\newcommand\blankpage{%
    \null
    \thispagestyle{empty}%
    \addtocounter{page}{-1}%
    \newpage}

\hypersetup{%2
    pdfborder = {0 0 0}
}
\usepackage{bm}
\usepackage[parfill]{parskip}

\newcommand*{\figref}[2][]{%
  \hyperref[{fig:#2}]{%
    {Figure~\ref*{fig:#2}}%
    \ifx\\#1\\%
    \else
      \,#1%
    \fi
  }%
}

\newcommand*{\tabref}[2][]{%
  \hyperref[{tab:#2}]{%
    {Table~\ref*{tab:#2}}%
    \ifx\\#1\\%
    \else
      \,#1%
    \fi
  }%
}

\newcommand*{\eqnref}[2][]{%
  \hyperref[{eqn:#2}]{%
    {Equation~\ref*{eqn:#2}}%
    \ifx\\#1\\%
    \else
      \,#1%
    \fi
  }%
}

\newcommand*{\chapref}[2][]{%
  \hyperref[{chap:#2}]{%
    {Chapter~\ref*{chap:#2}}%
    \ifx\\#1\\%
    \else
      \,#1%
    \fi
  }%
}

\usepackage{accents}
\newlength{\dhatheight}
\newcommand{\doublehat}[1]{%
    \settoheight{\dhatheight}{\ensuremath{\hat{#1}}}%
    \addtolength{\dhatheight}{-0.35ex}%
    \hat{\vphantom{\rule{1pt}{\dhatheight}}%
    \smash{\hat{#1}}}}
        
\DeclareMathAlphabet\mathbfcal{OMS}{cmsy}{b}{n}   

%\addtolength{\parskip}{1\parskip}
%\usepackage[square,comma,numbers,sort&compress]{natbib}
%
% Different fonts to try (uncomment only fontenc and one font at a time)
% (you may need to install these first)
%\usepackage[T1]{fontenc} %enable fontenc package if using one of the fonts below
%\usepackage[adobe-utopia]{mathdesign}
%\usepackage{tgschola}
%\usepackage{tgbonum}
%\usepackage{tgpagella}
%\usepackage{tgtermes}
%\usepackage{fourier}
%\usepackage{fouriernc}
%\usepackage{kmath,kerkis}
%\usepackage{kpfonts}
%\usepackage[urw-garamond]{mathdesign}
%\usepackage[bitstream-charter]{mathdesign}
%\usepackage[sc]{mathpazo}
%\usepackage{mathptmx}
%\usepackage[varg]{txfonts}

\hyphenation{dis-ser-ta-tion blue-print man-u-script pre-par-ing} %add hyphenation rules for words TeX doesn't know

%\renewcommand{\rightmark}{\scriptsize A University of California Davis\ldots \hfill Rev.~\#1.0 \quad Compiled: \currenttime, \today}
% a fancier running header that can be used with draftcls options

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                      %
%        DOCUMENT SETUP AND INFORMATION FOR PRELIMINARY PAGES          %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title          {Robust Image-Based Modeling and Simulation in Biomechanics}
%Exact title of your thesis. Indicate italics where necessary by underlining or using italics. Please capitalize the first letter of each word that would normally be capitalized in a title.

\author         {Omar Mohamed Hafez}
%Your full name as it appears on University records. Do not use initials.

\authordegrees  {B.S. Civil Engineering (University of California, Davis) 2009 \\
                 M.S. Civil Engineering (University of California, Davis) 2010 \\
                 M.S. Applied Mathematics (University of California, Davis) 2012}
%Indicate your previous degrees conferred.

\officialmajor  {CIVIL AND ENVIRONMENTAL ENGINEERING}
%This is your official major as it appears on your University records.

\graduateprogram{Civil and Environmental Engineering}
%This is your official graduate program name. Used for UMI abstract.

\degreeyear     {2018}
% Indicate the year in which your degree will be officially conferred.

\degreemonth    {March}
% Indicate the month in which your degree will be officially conferred. Used for UMI abstract.

\committee{Mark Rashid}{N. Sukumar}{John Bolander}{}{}
% These are your committee members. The command accepts up to five committee members so be sure to have five sets of braces, even if there are empties.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\copyrightyear{2018}
%\nocopyright

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\dedication{\textsl{To Mama, Baba, and Asmaa}}
\intentional{This page intentionally left blank}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\abstract{Image-based modeling and simulation has become an important analytic and predictive tool for patient-specific medical applications, including large-scale in-silico patient studies; optimized medical device design; and custom surgical guides and implants via additive manufacturing. The pipeline of steps for patient-specific modeling and simulation are: image acquisition, image segmentation, mesh generation, physics-based modeling and simulation, and finally clinical application. This research establishes a semi-automatic workflow for these steps, which includes a novel image-based meshing tool \textit{Shabaka}. The workflow is demonstrated in modeling the mechanics of a beating human heart based on magnetic resonance imaging (MRI) data.

Medical images are processed and segmented using thresholding, region-growing, and manual techniques. Watertight surface meshes are extracted from image masks using a novel Voronoi partitioning scheme. For scientific computing purposes, surface meshes are supplied either to tetrahedral meshing routines for conventional finite element approaches, or to a robust polyhedral mesh generation tool for a novel polyhedral finite element approach. The commercial polyhedral finite element code \textit{Celeris} is utilized, that retains most of the favorable properties of conventional finite element methods, while reducing the system size by up to an order of magnitude compared to conventional techniques for the same input surface.

In conjunction with the cardiac simulation code \textit{Cardioid} at Lawrence Livermore National Laboratory, the workflow is utilized to model the finite deformations of a beating heart. A quadratic tetrahedral mesh is generated from MRI data of the human heart ventricles. The constitutive law is comprised of an incompressible orthotropic hyperelastic stress response for the myocardium plus an electrical activation-dependent active stress for the muscle fibers. Muscle fiber orientations are generated using a rule-based approach. Electrical activation times are read from a coupled electrophysiology code. A lumped circulatory model is used to impose time-dependent ventricular volume constraints. Simulation results are presented. The same mechanics are also implemented for the polyhedral finite element mesh, and preliminary verification results are presented.

The tools developed and used for performing these image-based cardiac mechanics simulations make important improvements in the speed and robustness of image-based modeling techniques. As these tools continue to mature, so too does the promise for simulation to improve healthcare.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\acknowledgments{\noindent
Of the many lessons learned during this experience, one that tends to stick with me is that we are oftentimes more in control of our lives in the matters we think we are not, and less in control in the matters we think we are. The first part of that statement alludes to the hard work, self-reliance, and intrapersonal intelligence required to complete a PhD. The second part of that statement is an appreciation of just how much our successes in life depend on others. To that end, I humbly thank the following individuals, without whom I am utterly certain I would not have completed this degree:
 
I would first like to thank my advisor Professor Mark Rashid. I am forever grateful for his standard of excellence, precision of scientific communication, academic and emotional intelligence, and patience. I recall him telling me years ago ``you are not remembered for what you start, but for what you finish.'' I have kept those words of wisdom close to me ever since. There are few people over the course of my life I have learned more from than Professor Rashid.

My sincerest thanks to the Cardioid group for including me in their incredible work, in particular David Richards, Robert Blake, and Jean-Luc Fattebert at Lawrence Livermore National Laboratory, and Viatcheslav Gurev at IBM. Their mentorship and support provided me the opportunity to gain a much more complete perspective on my research, and opened opportunities for me I could have only dreamed of. I would also like to thank Andrew Baldwin and Alipasha Sadri at Celeris for their diligent work on my incessant feature requests in their software.

Many thanks to Professors Robert MacLeod, Jeffrey Weiss, and Ross Whitaker at the University of Utah for organizing and accepting me into their summer course on Image-Based Biomedical Modeling. It was easily one of the most intense, enjoyable, and memorable periods of time during my PhD.

I have many people to thank for providing me the financial support needed to complete this degree. I am forever indebted to Professor Mark Rashid and David Richards; the Krell Institute and the Computational Science Graduate Fellowship; Floyd and Mary Schwall; those in charge of the Eugene Cota Robles Fellowship; the UC Davis Department of Civil and Environmental Engineering; and Professors Sashi Kunnath, Amit Kanvinde, and John Bolander for giving me the means to accomplish my goals.
Financial support from DOE CSGF grant DE-FG02-97ER25308 is gratefully acknowledged. The work of LLNL authors was performed under the auspices of the U.S. Department of Energy by Lawrence Livermore National Laboratory under Contract DE-AC52-07NA27344.

Thank you to Professors John Bolander, N. Sukumar, and Yannis Dafalias for their guidance and mentorship over all these years dating back to my Master's Degree. Thank you to James St\"{o}lken and Ryan Vignes at Lawrence Livermore National Laboratory for their knowledge and wisdom during a period when I was first trying to fully understand what it meant to do good research.

A special thank you to the students I instructed in ENG 104. They injected an energy and enthusiasm in me in a way I think only a group of young students can. I would especially like to thank Bryan Zhao, who has inspired me to be a better teacher, a better researcher, and a better person.

Thank you to my office mates Steven Wopschall and Brian Giffin for giving me something to look forward to every day at the office. I will forever cherish the impact they have had on my life. Likewise, thank you to my roommate during the Image-Based Biomedical Modeling course, Brett Steineman, for whom I share the same sentiments.

Many thanks to Vincente Pericoli, Eric Chin, Subhajit Banerjee, and Alexandros Petalas for helping organize our student group, and thanks to Rose McCallen for her endless supply of enthusiasm, advice, and connections to help us get on our feet.

Thank you to my friends, among which there are too many to list here. Between lending an ear, providing advice, offering a couch to sleep on during my many visits to Livermore, or just bringing a smile to my face, I sincerely thank them all.

My love and thanks are extended to my mother, father, and brothers, who gave me the comfort and motivation to stay the course even through the most difficult times.

My deepest thanks to Asmaa Hassanein for being my rock, for helping me find my center again, and for believing in me.

Lastly, I thank God, for Whom I live my life.

I think we are all interconnected in ways we do not understand - I am grateful to be connected with all of these individuals, as well as those I have forgotten to mention, and those I did not even know I owed.
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Each chapter can be in its own file for easier editing and brought in with the \include command.
% Then use the \includeonly command to speed compilation when working on a particular chapter.
%%% \includeonly{chap1}

%\includeonly{chap7-conclusions}

\begin{document}

\newcommand{\bibfont}{\singlespacing}
% need this command to keep single spacing in the bibliography when using natbib

\bibliographystyle{ieeetr}
%many other bibliography styles are available (IEEEtran, mla, etc.). Use one appropriate for your field.

\makeintropages %Processes/produces the preliminary pages

\include{chap1-overview}
\include{chap2-medical-imaging-and-image-segmentation}
\include{chap3-image-based-meshing}
\include{chap4-numerical-methods}
\include{chap5-application-cardiac-mechanics}
\include{chap6-future-work}
\include{chap7-conclusions}
\phantomsection
\addcontentsline{toc}{chapter}{\hspace{6.5mm}References}
\bibliography{references}

\end{document} 